---
title: "Getting started with mwperm"
author: "The mwperm authors"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with mwperm}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(mwperm)
set.seed(1)
```

## What problem does this solve?

Many regressions have errors that are correlated within *several* overlapping
groups at once. The classic example is **dyadic data**: bilateral trade,
migration, or network flows indexed by an ordered pair $(i, j)$, where shocks
to country $i$ affect all of its partners and shocks to country $j$ affect all
of *its* partners. The error matrix is then exchangeable across rows and across
columns,
$$
\varepsilon_{ij} = \eta_i + \xi_j + u_{ij},
$$
and ordinary or even one-way cluster-robust standard errors are badly
anti-conservative.

The usual fix --- multi-way cluster-robust standard errors or the wild cluster
bootstrap --- relies on having *many* clusters and on regularity of the
covariates. When the number of clusters is modest, or a covariate is sparse or
heavy-tailed, those approximations can over-reject severely.

`mwperm` takes a different route: a **permutation test that is exact in finite
samples**. It tests
$$
H_0:\ \beta = b \qquad\text{in}\qquad
y_{ij} = x_{ij}^\top\gamma + d_{ij}^\top\beta + \varepsilon_{ij},
$$
assuming only that the error matrix is separately exchangeable across the
clustering dimensions. It makes **no assumption on the distribution of the
covariates** and needs no large-cluster asymptotics.

## Just one call: `mwperm()`

You do not need to know in advance which design applies. `mwperm()` inspects
the clustering structure — how many index dimensions, whether cells repeat,
whether the array is complete — and dispatches to the matching test,
forwarding every argument unchanged, so the result is identical to calling
the specific function yourself:

```{r unified}
data(trade_dyadic)
fit1 <- mwperm(y = "log_trade", d = "log_dist",
               x = c("log_gdp_i", "log_gdp_j"),
               index = c("importer", "exporter"),
               data = trade_dyadic, n_reps = 15, seed = 1)
```

To see the diagnosis without running anything, use `mwperm_check()`:

```{r check}
data(trade_panel)
mwperm_check(index = c("importer", "exporter", "year"), data = trade_panel)
```

Forks that the data itself can resolve (complete vs incomplete arrays,
repeated cells) are decided silently. Two forks hinge on an exchangeability
assumption the data *cannot* reveal, and there `mwperm()` defaults to the
choice that stays valid under the widest set of error processes and tells you
how to override it: a complete three-index array is treated as a **panel**
(time held fixed) unless you force `design = "threeway"`, because the panel
test remains valid when all three dimensions are exchangeable while the
converse fails; and repeated $(i, j)$ cells are treated as within-cell
replication (**layout**) with a warning, because they could equally be a time
dimension you forgot to pass — supply `time =` if so.

The rest of this vignette walks through the design-specific functions, which
remain fully supported for direct use.

## A dyadic example

The bundled `trade_dyadic` data set is a synthetic 40 × 40 gravity
cross-section. The coefficient on log distance was set to $-1$ in the
data-generating process.

```{r dyadic}
data(trade_dyadic)

fit <- with(trade_dyadic,
            mwperm_dyadic(y = log_trade, d = log_dist,
                          x   = cbind(log_gdp_i, log_gdp_j),
                          row = importer, col = exporter,
                          n_reps = 15, seed = 1))
fit
```

The point estimate is close to the truth, the p-value rejects $H_0:\beta = 0$,
and the confidence interval (obtained by **inverting** the test) excludes zero.

A placebo regressor that is independent of the outcome should *not* be flagged:

```{r placebo}
fit0 <- with(trade_dyadic,
             mwperm_dyadic(y = log_trade, d = placebo,
                           x   = cbind(log_dist, log_gdp_i, log_gdp_j),
                           row = importer, col = exporter,
                           n_reps = 15, seed = 2))
c(estimate = fit0$estimate, p_value = fit0$pvalue)
```

### Reading the output

`summary()` returns a tidy one-row-per-coefficient data frame, and `confint()`
extracts the inverted-test interval:

```{r accessors}
summary(fit)
confint(fit)
```

Because the test is *randomised* (it draws a random permutation group), running
it several times and aggregating by the median p-value is recommended; that is
what `n_reps` does. `plot(fit)` shows how stable the p-value is across those
replications.

## Resolution: why 20 clusters?

The smallest p-value the test can return is $1/(K+1)$, where $K+1$ is the order
of the permutation group and is capped by the number of clusters per dimension.
To form a 95% confidence interval you therefore need $1/(K+1)\le 0.05$, i.e. at
least **20 clusters** on the smaller dimension. With fewer clusters the test is
still valid, but `conf_int` is returned as `NULL` with an explanatory note:

```{r resolution}
small <- subset(trade_dyadic,
                as.integer(importer) <= 8 & as.integer(exporter) <= 8)
fits <- with(small,
             mwperm_dyadic(y = log_trade, d = log_dist,
                           x = cbind(log_gdp_i, log_gdp_j),
                           row = importer, col = exporter, seed = 1))
fits$conf_int      # NULL: 1/8 = 0.125 > 0.05
```

## Panel data

For longitudinal dyadic data, full three-way exchangeability is implausible
because errors are autocorrelated over time. `mwperm_panel()` instead assumes
exchangeability across the two cross-sectional dimensions *within each period*
and applies the **same** permutation in every period, which holds an arbitrary
common time trend $\zeta_t$ fixed. The `trade_panel` data set is a balanced
22 × 22 × 6 panel with a free-trade-agreement treatment whose true effect is
$0.5$.

```{r panel}
data(trade_panel)
fit_p <- with(trade_panel,
              mwperm_panel(y = log_trade, d = fta,
                           x   = cbind(log_gdp_i, log_gdp_j),
                           row = importer, col = exporter, time = year,
                           time_fe = TRUE, n_reps = 15, seed = 1))
fit_p
```

Keeping `time_fe = TRUE` (the default) removes the time trend from the
residuals, which de-biases the estimate and tightens the interval when
treatment timing is correlated with the period.

## Missing cells

When the array is incomplete --- for instance, if self-trade
($i = j$) is dropped --- general row/column permutations are no longer valid.
`mwperm_missing()` restricts the permutations to disjoint, fully observed
**bicliques** and pools the residual statistics across them, reporting how many
cells were retained.

```{r missing}
no_diag <- subset(trade_dyadic, importer != exporter)
fit_m <- with(no_diag,
              mwperm_missing(y = log_trade, d = log_dist,
                             x   = cbind(log_gdp_i, log_gdp_j),
                             row = importer, col = exporter,
                             min_block = 3, n_reps = 11, seed = 1))
fit_m
```

## Where to go next

- `?mwperm_dyadic`, `?mwperm_panel`, `?mwperm_missing` for the full argument
  lists.
- `?build_perm_set` for the permutation-group construction (Algorithm 1).
- The accompanying technical report works through these use cases in more
  depth, including a Monte-Carlo study of the test's size and power.

## References

Guo, W., Toulis, P. and Wang, Y. (2026). *Permutation inference under
multi-way clustering and missing data.* arXiv:2601.08610 [stat.ME].

Wen, K., Wang, T. and Wang, Y. (2025). Residual permutation test for regression
coefficient testing. *The Annals of Statistics* **53**(2), 724--748.
